Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-07-31 documentation/paper.tex
Reduced font size to 10pt; removed comments
Created stub for supervised learning section
Added section from interim report on performance analysis

\documentclass[12pt]{article}

% Use 1-inch margins even though the font is 12-point
\usepackage[margin=1in]{geometry}
\usepackage{textcomp}

@@

% Details on the data formats used as inputs for alpha/neutron discriminators
\section{Data Formats}

@@

>>>>

\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{textcomp}

@@

\section{Data Formats}

@@

\section{Performance Analysis}

The performance of neural networks relative to the Acoustic Parameter and the network used in the PICO-60 paper was done using the class-wise standard deviation. This is a variable defined as $C$ below, where $N$ and $A$ are the sets of outputs of the binary discriminator in question (Acoustic Parameter or a neural network), corresponding to the sets of neutrons and alpha particles respectively.

$S=std(N \cup A)$

$C=(std(N \div S) + std(A \div S)) \div 2$

The first step is to calculate the standard deviation S of the union of the two sets. This gives an indication of the scale of the overall distribution. When $N$ and $A$ are divided by $S$, they are normalized so that the standard deviation of their union is equal to 1. While the neural network’s outputs are bounded in the range of 0 to 1 with a sigmoid activation, Acoustic Parameter has a significantly wider range. Normalization of the union prevents this from creating a bias where Acoustic Parameter would produce a higher standard deviation with a similarly proportioned error.

The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually. This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are. Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off of those specific values, will produce a low class-wise standard deviation.

This performance metric is used rather than a simple accuracy score because it takes in to account not only whether the discriminator is correct or incorrect, but how confident it is. The intent is for a decisive discriminator, which produces a wide separation between the two classes, to be preferred over one that produces a nebulous cloud of outputs with a seemingly arbitrary decision boundary.

\section{Supervised Learning}

Supervised learning was used to directly replace Acoustic Parameter. It takes advantage of labeled sets of neutron and alpha events, which are used as training. Several different network configurations were applied to this task.

\subsection{Convolutional Neural Network for Raw Waveform Analysis}
